### 2013-06-12 paper/paper.tex
initial commit

\documentclass{sig-alternate}

\usepackage{listings}
\usepackage{hyperref}
\renewcommand{\topfraction}{0.9}
\renewcommand{\textfraction}{0.1}
\renewcommand{\floatpagefraction}{0.84}

\begin{document}
%
% --- Author Metadata here ---
\conferenceinfo{ICER 2013}{'13 San Diego, CA USA}
%\CopyrightYear{2007} % Allows default copyright year (20XX) to be over-ridden - IF NEED BE.
%\crdata{0-12345-67-8/90/01}  % Allows default copyright data (0-89791-88-6/97/05) to be over-ridden - IF NEED BE.
% --- End of Author Metadata ---

\title{An Open Platform for Managing Short Programming Exercises}

\numberofauthors{2}

% \alignauthor
% Andrei Papancea\\
% Jaime Spacco\\
%        \affaddr{Department of Computer Science}\\
%        \affaddr{Knox College}\\
%        \affaddr{Galesburg, IL 61401}\\
%        \email{\{apapance,jspacco\}@knox.edu}
% % 2nd. author
% \alignauthor
% David Hovemeyer\\
%        \affaddr{Kinsley Engineering Center}\\
%        \affaddr{York College of Pennsylvania}\\
%        \affaddr{York, PA 17403}\\
%        \email{dhovemey@ycp.edu}

\author{
\alignauthor
Author Name\\
Author Name\\
       \affaddr{Department of Computer Science}\\
       \affaddr{Small College \#1}\\
       \affaddr{City, State zipcode}\\
       \email{\{author1,author2\}@school.edu}
% 2nd. author
\alignauthor
Author Name\\
       \affaddr{Computer Science Department}\\
       \affaddr{Small College \#2}\\
       \affaddr{City, State zipcode}\\
       \email{author@school.edu}
}

%\date{30 July 1999}

\maketitle
\begin{abstract}

We have built CloudCoder, an open platform for creating, assigning, and sharing
short programming exercises for a variety of languages (we currently
support C/C++, Java, Python and Ruby, and can add others).  Like other
systems, CloudCoder is a web service that lets
students write code directly in their web browser, click the
``submit'' button, and receive immediate feedback.  Unlike other
systems, which tend to be closed, or commercial, or both, CloudCoder
is a completely open platform.  The code for the system is
open-source, and exercises written for CloudCoder are shared to a
central repository under a permissive Creative Commons license
(BY-SA).  Finally, CloudCoder collects detailed data that faculty can use for
educational research.

In this work, we report on successful pilot studies of CloudCoder at several
institutions.  We also describe on-going work to create a ``hint'' system
for CloudCoder, similar to what one would expect from an Intelligent
Tutoring System (ITS).

% Students access CloudCoder exercises 
% through their web browser, where they can edit code

% Importance of understanding code more deeply.

% Identify ``knowledge components''

% Identify when and where to insert hints

% (Maybe) use Euclidean distance to compute closeness of two solutions

% To generate hints:  What hints to generate, when to generate them

\end{abstract}

% A category with the (minimum) three required fields
\category{K.3.2}{Computers and Education}{Computer and Information
  Sciences Education}[Computer Science Education]

\terms{Human Factors, Measurement}

\keywords{CS1, novice programmers, data collection, student behavior,
  intelligent tutoring system}

\section{Introduction}

Recent years have seen an explosion of online services for learning
to program:  Udacity \cite{udacity}, CodeAcademy \cite{codeacademy},
ProgZoo \cite{progzoo}, 
Codingbat \cite{codingbat}, Coursera \cite{coursera}, Turingscraft
\cite{turingscraft}, Problets \cite{Kumar::2005:GPA:1163405.1163408}, CS Circles \cite{Pritchard:2013:CCI:2445196.2445370} and
several others.  These systems are very powerful, and show great promise to
improve introductory computer science.  Given all of these
very good systems, one may ask why we thought it a good career move
to build yet another system.  There are three main reasons:

%\begin{itemize}

%\item 
First, we wanted a system that an instructor can easily integrate into an existing traditional course.
  Many massive open online courses (MOOCs), such those available through Udacity and
  Coursera, were not designed for a traditional classroom and cannot be easily incorporated into an existing course.  
  The infrastructure provided by a MOOC is geared towards the course
  taught through the MOOC, and is not really configurable or customizable.

 % systems, such as Udacity, Coursera, and CodeAcademy, are
 %  massive open online courses (MOOCs) that cannot be easily
 %  incorporated into a traditional classroom setting.  MOOCs can be wonderfully
 %  enriching for students who take these courses.  However, because
 %  MOOCs offer an entire
 %  course, whatever online programming infrastructure they provide is part of that
 %  course.

%\item 
Second, we wanted an open, customizable system that focuses on short programming exercises that students
  can solve in their browser, and receive immediate feedback.  There are excellent systems that help
  faculty manage longer programming assignments
  \cite{Edwards:2008:WAG:1384271.1384371}, or homework and review questions
  \cite{turingscraft, Kumar:2005:GPA:1163405.1163408}.  Codingbat
  \cite{codingbat} is a fantastic free service that provides short
  programming exercises, but it is not easily customizable and does not provide
  access to all of the submissions submitted by students.
  We see CloudCoder as a complementary component to these resources.

% Some existing systems, such as Problets \cite{Kumar:2005:GPA:1163405.1163408},  and Turingscraft
%   \cite{turingscraft}, provide homework exercises that can be incorporated
%   such as Codingbat \cite{codingbat},
%  , can be integrated into an existing course.
%   Codingbat, while free to use and fantastically useful, does not give faculty access to a very
%   rich data set

%\item
Finally, we wanted exercises that could be freely modified and
redistributed.  Much of the excellent content available through MOOCs,
or the increasing amount of digital content available through textbook publishers, 
cannot be modified or re-used outside of those contexts.  This can be a significant
impediment to researchers trying to study the difficulty level of
various exercises, or to build intelligent tutoring systems (ITSs)
that require a large bank of questions and their solutions.

%\end{itemize}

% There are two reasons.  First, many of
% these systems, such are Udacity, Coursera, and CodeAcademy, are
% massive open online courses (MOOCs).  MOOCs typically provide
% supporting infrastructure that helps students learn to code in their
% browsers, and the course material offered by a MOOC can be a wonderfully enriching
% experience for someone who enrolls in the course.  However, it is not
% clear how faculty who teach more traditional courses can
% integrate any of the infrastructure provided by a MOOC into
% their curriculum.  Second, 
% the systems that could be easily integrated into an existing course, such as
% Codingbat or Turingscraft, are
% are either commercial, or closed, or both.  To our knowledge, none of
% these systems readily support faculty who want to 
% modify the code, host their own installation, extract data about
% student behaviors beyond the basic correctness of their submissions,
% or redistribute the programming exercises and assignments.

We believe the academic community should have a fully open alternative
to the closed and/or commercial systems that have exploded onto the CS
Education scene.  Faculty can integrate CloudCoder into an existing course with minimal overhead,
can write their own content, use and modify content contributed by
others, customize the system, and collect data for educational studies.

\section{CloudCoder}
We have designed CloudCoder to be an open platform for managing short
programming exercises.  The CloudCoder platform has four design
goals:  To be an open-platform for faculty to assign short programming
exercises, to build a collection of redistributable programming
exercises, to collect data about student behaviors as they solve
exercises, and to explore ways of curating the collected data.

% \begin{itemize}
% \item {\em To provide a web-based platform for faculty to assign short
%   programming exercises.}

% There are platforms such as Web-CAT \cite{Edwards:2008:WAG:1384271.1384371},
% 
%  TODO Insert Marmoset reference!
%
 % and the many in-house systems developed at universities, for managing longer programming
 %  assignments, and there are commercial systems such as Turingscraft
 %  for assigning online homework \cite{turingscraft}.  However, there
  % are currently no open, flexible systems that serve a similar
  % purpose.  All of the code is open-source and available at Github:

% \item {\em To crowd-source the construction of a repository of programming
%   exercises.}  
% \item {\em To collect data about student work on short
%   programming exercises.}  There are many systems that collect data
%   about novice programmers \cite{Utting:2012:WDG:2361276.2361278,
%     Jadud:2006:MTE:1151588.1151600, Norris:2008:CCQ:1384271.1384284,
%     Spacco:2006:EMD:1140124.1140131}. % TODO: Cite French paper on Eclipse plugin
%   However, these systems typically target longer programming
%   assignments that students spend a week or two working on.
%   CloudCoder target very short exercises, and logs data at the
%   keystroke level.  We envision faculty using CloudCoder as a platform
%   for educational studies that require studying fine-grained trace data.
% \end{itemize}

\subsection{An Open Platform}

CloudCoder is an open platform for managing short programming
exercises.  Each exercise can be solved with one method or one short
program, and could take only a few minutes to solve, or could require
or an hour or two.  The system is not designed to support longer
programming assignments, such as those found at the {\em Nifty
  Assignments} session at SIGCSE.

Figure~\ref{screenshot} shows a screenshot of a CloudCoder exercise.
The editor is derived from the Ace editor \cite{ace}, and supports
syntax high-lighting for most programming languages, and
auto-indentation.  
A submit button runs the code against a series of
test cases, and feedback is provided directly in the browser.
CloudCoder captures anything the code prints to standard out or
standard err to help students with debugging.

\begin{figure*}[ht]
\centering
\includegraphics[width=180mm]{images/screenshot.png}
\caption{Screenshot of the main view in CloudCoder.}
\label{screenshot}
\end{figure*}

CloudCoder is available under the GNU Affero open-source license.
Binary distributions are available through our website (anonymized),
and the code is available on Github:

\vspace*{3mm} 
%\url{https://github.com/cloudcoderdotorg/CloudCoder}
(link anonymized)
\vspace*{3mm}

% The editor window

\subsection{Permissively-Licensed Exercises}

Writing good homework exercises is difficult and
time-consuming, and even the best textbooks are limited in the
number and quality of practice exercises they can provide, to say
nothing of a textbook's inability to provide timely feedback to
students.
Sharing exercises between faculty members is a much
more efficient way to increase the number and quality of exercises at
our disposal.

Faculty who use CloudCoder can write their own exercises using our in-browser
authoring tools, or can import exercises written by other faculty
from a central repository at:

\vspace*{3mm} 
% \url{https://cloudcoder.org/repo}
(link anonymized)
\vspace*{3mm}

Currently we have over 50 Java and 60 C/C++ exercises authored by 5
different users.  By default,
any new exercises written by faculty will be shared to the 
central repository under a permissive Creative Commons license (BY-SA)
that allows redistribution provided that the author is credited and
redistribution is under the same license.  
(Note that the repository shows the exercise description, but not the
test cases or the solution, so an unscrupulous student cannot dig through the
repository for solutions.  Importing exercises requires an account at
the repository, which can be used to discourage this sort of
undesirable behavior.).

CloudCoder supports two types of exercises:  {\em method} or {\em function-based}
exercises, and {\em whole-program} exercises.  Function based exercises look
a lot like Codingbat exercises:  Students are given a description of
what the method should do, and a couple of example input/output pairs
similar to a unit test.  Writing a new function-based exercise
requires a description, some ``starter code'' (typically the method
header, although exercises can provide as much or as little code as
desired), and a series of inputs and their corresponding outputs.
Test-cases can be marked as ``secret'' so that students don't see their
outputs to disincentivize students from writing a special case to
recognize each input.

On the other hand, whole-program exercises
have a description, but read inputs from
standard-in and write to standard-out.  CloudCoder uses a
simplified regular expression language to check student outputs for
correctness.

\subsection{Building and Testing Submissions}

CloudCoder currently supports exercises for C/C++, Java, Python and
Ruby.  Building and testing takes place in a separate process which
can be run on a separate machine for extra security, or on multiple
machines for scalability.

We run Java code relatively securely under a Security Manager, and 
use Jython and JRuby to compile Python and Ruby to Java
bytecode, which we also run in a Security-Managed sandbox.  We
currently support C/C++ less securely, although 
recent work by Malan \cite{Malan:2013:CSS:2445196.2445242} on a
``sandbox'' for running untrusted student code can be leveraged to improve our security.

\subsection{Data Collection and Curation}

CloudCoder logs keystrokes in the online editor window, and 
stores all submissions and test outcomes in a databases.
We hope that the combination of exercises and data will give faculty
insights into the difficulty level of different exercises, keep a
record of common error patterns, and enable the development of
innovative data analyses.

One question is how to manage the data collected.  
Recent work on web-scale data collection by the BlueJ team
\cite{Utting:2012:WDG:2361276.2361278} lays out a number of important
questions about programming data, such as who should have access to the data, and what
mechanism should be used to access the data.  Sanders
et. al. \cite{Sanders:2008:DSE:1404520.1404534} reviewed a number of
the challenges involved in sharing empirical education data.

An alternative repository for the type of data collected by CloudCoder
is the Pittsburgh Science and Learning Center (PSLC) Datashop, which
is a repository of data collected by by intelligent tutoring
systems (ITSs).  By its nature, ITS data is much simpler than program
snapshot data, so snapshots cannot be added without some changes.
The Datashop is interested in storing program
snapshot data to make it easier for researchers in the ITS, educational data
mining, and CS-ed communities to work on programming data.

Anonymized data that has been cleared by the proper Institutional
Review Boards is contributed to the Datashop, which
makes the data available to other researchers under whatever
restrictions were put in place by the overseeing IRBs.  The Datashop also
provides a suite of analysis and visualization tools for better making
sense of the ITS data.

We are working with members of the Datashop to establish standards for
program snapshot datasets, and to begin developing tools and
techniques that can help analyze this very different dataset.  This is
a long-term project that is in its infancy.

% The Datashop stores traces from ITSs, which are typically made up of
% multiple choice questions, and merely need to store a student's
% answer, and whether they asked for a hint.

% We have been in contact with the the Pittsburgh Science and Learning
% Center (PSLC) Datashop.  The Datashop is a repository for

% The data collected by CloudCoder is simpler than the data collected by
% the BlueJ team.

%Standard storage formats, visualization tools, standard analyses.

\subsection{Hosting CloudCoder}
We recognize that ease of installation is a critical factor for
adoption of a new pedagogical
technology for busy
faculty who may only have a few minutes to evaluate a new system.  We
are committed to making the installation process as simple and painless as 
possible.

Thus far each institution that has adopted CloudCoder has
provided a Linux server that we configured for them.  This has made
adoption very easy, but is unlikely to scale up as we acquire more
users.  We have asked Amazon for Amazon Web Services credits to
provide CloudCoder ``in the cloud'' to adopting institutions, and we
hope to add support for Google App Engine this summer.  Ease of
adoption is critical for CloudCoder.

\subsection{Pilot Studies}

We piloted CloudCoder at several institutions during the 2012-2013
academic year.  Table~\ref{tab:courses} shows a breakdown of the
courses.  We tried to use CloudCoder at Auckland in Fall 2012 for a
course with 700 students, but the free-tier of Amazon EC2 did not
scale up well enough, so we only recorded partial data.  We have made
significant improvements since then and plan to deploy CloudCoder for
the same large course in Fall 2013.

\begin{table*}
\centering
\begin{tabular}{| l | l | l | l |}
\hline
Institution & Course & Language & Enrollment\\
\hline
Canisius College & CS-1 & C/C++ & 30\\
Small College \#2 & CS-1 & C/C++ & 120\\
Small College \#1 & CS-1 & Java & 23\\
University of Auckland & Programming for Engineers & C & 30\\
\hline
\end{tabular}
\caption{CloudCoder pilot studies.}
\label{tab:courses}
\end{table*}

In the Winter 2013 term at Small College \#1, students taking CS-1 used {\tt CloudCoder}
%
% TODO: \cite{Hovemeyer:2013:CBC:2445196.2445451}.  
% 
We collected over 6,000 submissions
for 32 exercises by 23 students.  We expect to have a much larger
dataset from 120 students at 
Small College \#2
%York College 
once the Spring 2013 semester ends.  Our analysis of these datasets is
very preliminary at this point, but work is ongoing.

\section{Discussion}

We could use community input into the CloudCoder project.  Specifically:

\begin{itemize}
\item Who would be interested in adopting this system?  If you would
  consider trying out the system if we added a particular feature,
  what would that feature be?  We would like
  to expand our pilot group.
\item What studies could or should be done with this platform?  And
  would you like to be involved with these studies?
\item What studies can be done with the existing data we have
  collected?  We do not yet have well-developed tools for data
  analysis, but one data set has been anonymized and can be
  distributed to other researchers interested in our data.
\end{itemize}

\section{Related Work}

Jadud's pioneering early work on novice compilation behavior
\cite{Jadud:2006:MTE:1151588.1151600}, along with follow-up work by
others \cite{Norris:2008:CCQ:1384271.1384284} provides a starting point for
work with program snapshots.

Denny et. al. have studied novice syntax errors
\cite{Denny:2012:SEE:2325296.2325318} using data collected from the CodeWrite system.
CodeWrite is similar to CloudCoder, except that students create
exercises and test cases for each other to solve, and are evaluated on
the quality of their practice exercises.

Lane et. al. describe Intention-Based Scoring (IBS) \cite{Lane:2005:ISA:1047344.1047471},
an approach to building an ITS for programming whereby human
judges to score or code {\em charettes}, or assignments students work on in a
timed, closed-lab environment.  IBS derives from prior work on
identifying ``bugs'' in online protocols, and as such attempts to
identify places where students have a ``bug'' in their development
process.  This approach, while labor-intensive, shows some promise in
developing an ITS for programming assignments.

Zingaro et. al. built the Python Classroom Response System
\cite{Zingaro:2013:FCP:2445196.2445369}, a system similar to
CloudCoder, except that it focuses on supporting in-class programming
questions for a class that uses a {\em Peer Instruction} pedagogical approach.

Pritchard et. al. present CS Circles
\cite{Pritchard:2013:CCI:2445196.2445370}, an interactive on-line
Python curriculum that embeds short questions, some of which involve
programming, directly into the text.  This sort of interactive text
shows great promise for developing open, interactive courses.

The BlueJ team presents a compelling case for web-scale data
collection \cite{Utting:2012:WDG:2361276.2361278}.  Their system is
more ambitious than CloudCoder, but the data collected by CloudCoder
should be complementary.

\section{Future Work: Towards an Intelligent Tutoring System}

One of the long-term goals for CloudCoder is to use the data to
develop an Intelligent Tutoring System for programming assignments.
A web-based programming ITS could be used to support high school
AP courses in school districts that may lack computer science
teachers, or to help improve introductory CS courses by reducing
student frustration.

To develop an ITS, we need a large supply of exercises; we hope
CloudCoder's permissible licensing will help on this front.  We also
need to provide students with ``hints'' when they are stuck.  To
provide hints, we can either write the hints ourselves, which is
labor-intensive, or generate them automatically by mining them out of
the dataset of prior submissions.  Obviously this second approach is
a much preferred method of creating hints.  However, identifying hints
automatically means that we need to match a student's current program
state with the most similar program state we've seen previously, 
figure out what successful next step the previous student took, and
then summarize this information into a hint.  This is a very
challenging problem.

There are currently two promising lines of research
trying to address the issue of identifying the most similar previous submissions.  Rivers and Koedinger \cite{rivers-its2012} have developed a
system that computes a {\em canonicalized} version of each submission
by compiling to an Abstract Syntax Tree (AST) and then applying a
series of compiler transformations.  This work is ongoing, but
preliminary results are encouraging.
On the other hand, Jin et. al. \cite{Jin:2012:PRA:2345840.2345889, jin-kdd2011} perform
an inter-procedural data flow analysis that tracks the values stored
in different variables in an effort to identify variables that are
used in the same way in each submission.  They also show promising preliminary
results for their approach.

\subsection{Key Features:  Mining Hints from the AST}

Our approach, which is very preliminary, ongoing research,  is also based on an analysis of
the AST.
Our hypothesis is that many novice-level exercises
will have solutions that require certain structural
{\em key features} in the code that we can detect automatically by
comparing the AST of the student's code with the ASTs of known correct
solutions.

For example, Figures~\ref{fig:counta} and \ref{fig:countab} show two
CloudCoder exercises that were assigned during the Winter 2013 term in
CS-1 at 
Small College \#1
%Knox College.

\begin{figure}[h]
{\tt Write a static method called countA that takes a String as a
parameter and returns the number of times that 'a' occurs in the String. }
\caption{CountA:  An example CloudCoder exercise.}
\label{fig:counta}
\end{figure}

\begin{figure}[h]
{\tt Write a static method called countAB that takes a String as a
parameter and returns the number of times that ab occurs (i.e. an 'a'
immediately followed by 'b'). }
\caption{CountAB: Another example CloudCoder exercise.}
\label{fig:countab}
\end{figure}

First let us consider the {\tt CountA} exercise from
Figure~\ref{fig:counta}.  As is typical in computer science, students have a lot of flexibility
in how they solve this problem:  with a for loop or with a while loop, using the {\tt charAt()}
method or the {\tt substring()} method or other String methods, and so on.  However, in
practice, exercises are assigned in a context with other
exercises, in-class examples, and course topics, which tends to limit
the scope of possible student solutions, and to push students towards
certain solutions and away from others.  For example, no student solved this
exercise with a while loop, and all but one used {\tt charAt()}
rather than other String operations.  In fact, most correct
solutions looked something like Figure~\ref{fig:counta-solution}.

\begin{figure}[h]
\begin{lstlisting}
public static int countAB(String str) {
  int count=0;
  for (int i=0; i<str.length();i++) {
    if (str.charAt(i)=='a') {
      count++;
    }
  }
  return count;
}
\end{lstlisting}
\caption{Typical solution for Exercise CountA depicted in Figure~\ref{fig:counta}.}
\label{fig:counta-solution}
\end{figure}

\begin{figure}
\includegraphics[width=90mm]{images/countA.pdf}
\caption{Slightly simplified AST of the code in Figure~\ref{fig:counta-solution}}
\label{fig:counta-ast}
\end{figure}

Compiling the method from Figure~\ref{fig:counta-solution} to an
Abstract Syntax Tree (AST) produces a tree that looks something like
Figure~\ref{fig:counta-ast} (the AST has been simplified slightly by
collapsing nodes that are unimportant for this example).  We can identify
several features of the AST that must be present to solve the
exercise, such as a for loop, an if-statement inside the loop, and
the use of a String method, typically {\tt charAt()}, to examine
characters inside the loop.  We can also ensure that the components of
the for loop are correct, i.e. that the loop counter starts at 0, that
the loop goes up to the length of the String, and so on.  Assuming
that student code compiles, all of these necessary features can be
extracted from the AST.

Now let's consider the {\tt CountAB} exercise depicted in
Figure~\ref{fig:countab}.  Again, there were many correct solutions;
however, many correct solutions looked something like
Figure~\ref{fig:countab-solution}.

\begin{figure}[h]
\begin{lstlisting}
public static int countAB(String str) {
  int count=0;
  for (int i=0; i<str.length()-1;i++) {
    if (str.charAt(i)=='a' && 
        str.charAt(i+1)=='b') {
      count++;
    }
  }
  return count;
}
\end{lstlisting}
\caption{Typical solution for Exercise countAB depicted in Figure~\ref{fig:countab}.}
\label{fig:countab-solution}
\end{figure}

Compiling the method in Figure~\ref{fig:countab-solution} into an
abstract syntax tree (AST) produces a tree that looks like
Figure~\ref{fig:countab-ast} (this result has been slightly simplified
from the full output from the Eclipse JDT compiler).

\begin{figure}[h]
\includegraphics[width=90mm]{images/forloop-highlight.png}
\caption{Slightly simplified AST of the code in Figure~\ref{fig:countab-solution}}
\label{fig:countab-ast}
\end{figure}

Note the two circled sub-trees, or {\em features}, in the AST depicted
in Figure~\ref{fig:countab-ast}.  Feature \#1, the {\em
  length-1} feature, shows the loop iterating
until the length of the String minus one.  The for loop requires a
stop condition that is different
than the loop stop condition used to solve the {\tt CountA} exercise from
Figure~\ref{fig:counta}.  Crucially, once we know students are
using a for loop to solve this exercise (which we can also
extract from the AST), {\em we know exactly where in the AST to
  look for the length-1 feature}.  If a submission has all the
features except the {\em length-1} feature, we have a pretty good idea
that the student hasn't figured out to change the stop condition of the
for loop.  We could use the lack of features present in correct
solutions to give us an idea of what things the student is missing.

\subsection{Limits of Key Features Approach}

Our approach of identifying ``key features'' has a number of
weaknesses and limitations.  First, like other ITS hint-systems, it
requires a reliable system for identifying the nearest previous successful
submission to a student's current program state.  Work in this area is
ongoing \cite{rivers-its2012,Jin:2012:PRA:2345840.2345889}.

Second, features are based purely on the structure of the
code, and are treated independently of one another, which is a
simplification that throws out a lot of relevant contextual information.  For
example, a student submission may have two for loops, one of which has
an if statement nested inside it; our approach cannot determine which
loop has the if statement inside it.

Third, we do not perform any data-flow analysis (at least not yet), so
we cannot identify if particular values are stored in variables.

Next, we require compilable code, which can be a problem for the many
students who struggle to get their code to compile
\cite{Denny:2012:SEE:2325296.2325318}.

Finally, while we focused mainly on exercises that required for loops
to solve, we also assigned a number of exercises that
use if statements and no loops.  For these exercises, we can figure
out when students don't have 
enough structure, such as when they don't have nested if statements
for an exercise where most people who solves the exercise used nested if statements.  However,
we cannot yet easily determine when students have the wrong boundary
conditions in an if statement.

Much more work is needed in this area.  We hope that 
CloudCoder can help out by providing a rich data set for post-hoc
studies, as well as a platform for testing out a hint-system.

\bibliographystyle{abbrv}
\bibliography{paper}

%\appendix
%Appendix A
\section{Acknowledgements}

Anonymized.

\end{document}