### 2013-07-23 paper/talk.tex
motivation slide

\documentclass{beamer}
\usepackage{graphicx}
\usetheme{Boadilla}

\begin{document}

\title[ICER 2013]{An Open System for Managing Short Programming Exercises}
\author[Papancea et.\ al.]{Andrei Papancea, Jaime Spacco, and David Hovemeyer${}^{\dagger}$}
\institute[Knox, YCP]{
\begin{minipage}[t]{1.1in}
\begin{center}
{{\ }\\
Knox College}
\vskip .1in
\includegraphics[height=.8in]{images/KnoxLogo}
\end{center}
\end{minipage}
\begin{minipage}[t]{1.1in}
\begin{center}
{${}^{\dagger}$York College\\
of Pennsylvania}
\vskip .1in
\includegraphics[height=.8in]{images/YCPLogo}
\end{center}
\end{minipage}
}
\date[August, 2013]{August 12, 2013}

\begin{frame}[plain]
  \titlepage
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%% Outline %%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{Outline}

\begin{itemize}
  \item Background
  \item Description of CloudCoder
  \item Preliminary Experiences
  \item Questions
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%% Background %%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{Background}

\begin{itemize}
  \item Many students struggle in introductory programming courses
  \item Mastery of basic concepts and techniques is essential for success
  \item Short programming exercises:
  \begin{itemize}
    \item Students write or complete a short (5-15 line) program
    \item Correctness is judged with a series of test cases
    \item Student receives immediate feedback about results of tests
  \end{itemize}
  \item Exercises are useful for:
  \begin{itemize}
    \item Skills development
    \item Self-tests to accompany reading
    \item Extra practice
  \end{itemize}
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%% Why another system? %%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{Why another system?}

\begin{itemize}
  \item There are lots of programming exercise systems already:
  \begin{itemize}
    \item Free: CodingBat, Codecademy
    \item Commercial: CodeLab (TuringsCraft), MyProgrammingLab (Pearson)
    \item {\em Do we really need another one?}
  \end{itemize}
  \item We did not know of one that
  \begin{itemize}
    \item Was open source
    \item Encouraged users to create exercises
    \item Allowed easy sharing of exercises
    \item Supported a range of languages (including C/C++)
  \end{itemize}
\end{itemize}

\end{frame}

\end{document}